---
title: "Onshore wind and the likelihood of planning acceptance: learning from a Great Britain context"
author: "Michael Harper, Ben Anderson, Patrick James, AbuBakr Bahaj"
output: 
  bookdown::pdf_book:
    number_sections: true
    keep_tex: true
    toc: FALSE
bibliography: [bibliography.bib, software.bib]
link-citations: true
papersize: a4
csl: elsevier-with-titles.csl
header-includes:
    - \usepackage{setspace}\doublespacing
---

```{r setup, include=FALSE}
# devtools::install_local("WindAnalysis")
library(WindAnalysis)
library(tidyverse)

# Specify knitr options
knitr::opts_chunk$set(echo = FALSE, 
                      out.width = "50%",
                      tidy = TRUE,
                      fig.align = "center", 
                      out.width = "8.8cm",
                      fig.pos = "h",
                      cache = F)


# Set the global theme for figures
ggplot2::theme_set(theme_Publication(base_family = "serif"))
```


```{r loadCache, include=FALSE}
# Run the full analysis if the cache does not exist
if(!file.exists("_code/outputs/codeOutputs.RData")){
  
  setwd("_code/")
 bookdown::render_book("AnalysisComplete.Rmd")

 }

# Load the cached results
load("_code/outputs/codeOutputs.RData")
```


## Abstract {-}

Geospatial modelling is extensively used to identify suitable sites for the installation of onshore wind turbines. However, there are concerns that such approaches do not accurately consider the social issues surrounding such projects, resulting in large numbers of projects subsequently being rejected at planning permission. Using the location of 1691 historic wind turbine planning applications in Great Britain, this paper explores whether the planning success of proposed wind turbine projects can be predicted using a range of geospatial, social and political parameters. The results indicate that the size of the project, percentage of the local population with high levels of qualifications, the average age, and the proximity to existing wind turbines are key influences affecting planning approval. The paper demonstrates that quantitatively linking regional social and political data enhances the assessment of the planning outcome of wind turbines, and highlights that geospatial parameters are in themselves limited in identifying the suitability of sites. This suggests that existing policy is...

## Keywords {-}
Onshore Wind, Logistic Regression, Planning, Demographics, Great Britain, GIS

**Word Count**: `r HarpR::countWordsRmd("Paper.Rmd")[[1]]`

#	Introduction

Increased environmental concern and issues surrounding security of energy supply have led to a global drive to develop renewable energy systems. Over \$40 billion is invested annually within the European Union, with this figure expected to exceed \$60 billion by 2020 [@UNEP2016]. 

Whilst many renewable energy technologies are available, onshore wind is is one of the most established technologies and offers one of the least-cost options for renewable energy supply. For example, the cost projections for new onshore wind projects in the United Kingdom in 2020 are projected to be between £47-76/MWh, a price which competes with conventional fossil-fuel technologies [@DBIES2016]. This economic viability is coupled with a high resource availability, with the UK and many other European countries having a large exploitable wind resource [@EuropeanEnvironmentAgency2009].

Despite the strengths of onshore wind energy, development of the technology is restricted as there are challenges for proposed projects to receive planning permission. Proposals often face local opposition, with visual impact, noise, site access and ecological impacts often being cited as reasons for objection [@Wolsink2000; @Langer2016]. These planning challenges are particularly evident in the United Kingdom, where 52% of onshore wind projects are refused permission or are abandoned by the developer [@DECC2016].  As highlighted in Figure \@ref(fig:acceptanceRates), this rate is signficantly higher than for other renewable energy technologies in the UK.

(ref:acceptanceRates) Acceptance rates of renewable energy projects within the UK between 1991 and 2017. The Analysis only considers technologies which have had more than 50 planning applications.

```{r acceptanceRates, fig.cap="(ref:acceptanceRates)", cache = TRUE}

# A summary table was made to convert the summary status
planning_summary <- read.table(text = "
'Status'	                    'Status Summary'
'Application Approved'	      'Approved'
'Application Refused'	        'Refused/Abandoned'
'Application Submitted'	      'Submitted'
'Application Withdrawn'	      'Refused/Abandoned'
'Connection Applied For'	    'Submitted'
'No Application Made'	        'Refused/Abandoned'
'No Application'              'Required	Approved'
'Abandoned'	                  'Refused/Abandoned'
'Awaiting Construction'	      'Approved'
'Decommissioned'	            'Approved'
'Operational'	                'Approved'
'Planning Permission Expired'	'Approved'
'Under Construction'	        'Approved'
", header = TRUE, stringsAsFactors = FALSE)

REPD_count <-
read.csv(file = "_data/Public_Database_-_Feb_2018.csv", check.names = TRUE, as.is = TRUE) %>%
  filter(Country != "Northern Ireland") %>%
  merge(planning_summary, by.x = "Development.Status..short.", by.y = "Status") %>%
  dplyr::select(Technology.Type, Status.Summary) %>%
  set_names(c("Tech", "Status")) %>%
  table() %>%
  as.data.frame() %>%
  reshape2::dcast(Tech ~ Status, value.var = "Freq") %>%
  mutate(n = rowSums(.[2:4]),
         App_perc = `Approved`/n)

# Convert to long for plotting
REPD_count_long <-
  REPD_count %>%
  reshape2::melt(id = c("Tech", "n", "App_perc")) %>%
  set_names(c("Technology", "Group_Total", "Perc_Approv", "Status", "Count")) %>%
  mutate(Status = factor(Status, levels = c("Approved", "Submitted", "Refused/Abandoned")),
         Technology = factor(Technology, levels = rev(REPD_count$Tech[order(REPD_count$App_perc)])),
         percent = Count/Group_Total) %>%
  filter(Group_Total > 50)

# Plot the results
ggplot(REPD_count_long, aes(x = Technology, y = Count, fill = Status)) +
  geom_bar(position = "fill", stat = "identity", alpha = 0.9, width = 0.7, color = "grey20") +
  geom_text(aes(y = percent, label=ifelse(percent >= 0.07, paste0(sprintf("%.0f", percent*100),"%"),"")),
                position=position_stack(vjust=0.5), colour="grey20", size = 3) +
  coord_flip() +
  theme(legend.position = "top", aspect.ratio = 1,
        plot.margin = unit(c(1,3,1,1), "lines")) +
  scale_fill_manual(values = c("#91cf60", "#fee08b", "#fc8d59")) +
  scale_y_continuous(labels = scales::percent, expand = c(0,0)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 20)) +
  labs(x = "", y = "") +
  guides(fill = guide_legend(reverse = TRUE))
```

Whilst planning has been a persistent issue for wind turbines, recent changes in legislation have severly impacted the development of onshore wind. Until June 2015, the planning decision for projects greater than 50MW was controlled at a national level. However, this provision was removed by the *Energy Act 2016* and *Infrastructure Planning Order 2016*,  which provided local authorities with the final say for all onshore wind energy projects [@Smith2016]. In addition, these wind turbines are only permitted in sites which have been identified within neighbourhood development plans. These changes have effectively enabled local communities to be in a position to block the development of wind turbines in their area.

```{r}
turbine_dates <- read.csv(file = "_data/Public_Database_-_Feb_2018.csv", check.names = TRUE, as.is = TRUE) %>%
  mutate(date = lubridate::dmy(`Planning.Application.Submitted`),
         year = lubridate::year(date),
         month = lubridate::floor_date(date, unit = "months")) %>%
  filter(Country != "Northern Ireland") %>%
  merge(planning_summary, by.x = "Development.Status..short.", by.y = "Status") %>%
  mutate(Status.Summary = factor(Status.Summary, levels = c("Approved", "Submitted", "Refused/Abandoned"))) %>%
  filter(!is.na(year)) %>%
  filter(year != 2018)

turbines_count_month <- 
  turbine_dates %>%
  dplyr::group_by(month) %>%
  dplyr::summarise(sum = n()) %>%
  group_by(month) %>%
  mutate(freq = sum / sum(sum)) %>%
  as.data.frame() %>%
  filter(lubridate::year(month) %in% 2013:2018)

turbines_count_year <- 
  turbine_dates %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(sum = n()) %>%
  group_by(year) %>%
  mutate(freq = sum / sum(sum)) %>%
  as.data.frame()

applications_before <- turbines_count_month %>%
  filter(month < as.Date("2015-07-18") & month > as.Date("2014-07-18")) %>%
  summarise(count = sum(sum)) %>%
  as.numeric()

applications_after <- turbines_count_month %>%
  filter(month < as.Date("2016-07-18") & month > as.Date("2015-07-18")) %>%
  summarise(count = sum(sum)) %>%
  as.numeric()

applications_after2 <- turbines_count_month %>%
  filter(month < as.Date("2017-07-18") & month > as.Date("2016-07-18")) %>%
  summarise(count = sum(sum)) %>%
  as.numeric()

```

In addition to changes in planning law, the development of onshore wind has been further restricted by changes to the financial mechanisms used to support low-carbon energy in the UK. Onshore wind projects were removed from the Renewable Obligation scheme on 1st April 2016 [@], preventing projects from bidding in the upcoming £557 million (date: ) in subsidies [@].

As a result these planning and financial changes, there has been a dramatic reduction in the development of onshore wind within the UK. In June 2015, the month the changes in planning were implemented, there were 133 planning applications for onshore wind, a record high for a single month [@DECC2016]. In contrast, for the entire year of 2017, there were only 52 planning applications made, representing only 6% of the 2015 total. This reduction in planning is highlighted in Figure \@ref(fig:numberApplications).

(ref:numberApplications) Number of turbine planning applications submitted in the United Kingdom per month between January 2013 and December 2017.

```{r numberApplications, fig.cap="(ref:numberApplications)"}

labels <- data.frame(x = c(as.Date("2015-07-01"), as.Date("2016-04-01")), y = c(150, 150), label = c("One","Two"))

ggplot() +
  geom_col(data = turbines_count_month, aes(x = month, y = sum), 
           fill = "#70AFD1", colour = "black", size = 0.2) +
  labs(y = "Number of Applications per month",
       x = "Year")  +
  theme(aspect.ratio = 1,
        plot.caption = element_text(hjust = 0)) +
  scale_y_continuous(limits = c(0, 150)) +
  scale_x_date(date_breaks = "1 years", date_labels = "%Y") +
  geom_vline(xintercept = c(as.Date("2015-07-01"), as.Date("2016-04-01")), linetype  = "longdash") +
  annotate(geom = "label",
           x = c(as.Date("2015-07-01"), as.Date("2016-04-01")),
           y = c(150, 150),
           label = c("1","2")) +
  labs(caption = "1: Change in planning law to provide local control \n2: Removal of financial support for onshore wind")
```

Whilst the predicted reduction of cost in onshore wind may overcome the financial barriers in the long term, the planning changes have act as a major barrier to increased. This therefore highlights the importance of considering local communities within the proposals of onshore wind projects and understanding characteristics which may lead to project rejection. The UK is not alone in experiencing in encountering opposition to wind turbine projects [@Langer2016], but is perhaps unique in how policy has been restructured to restrict its development.

This paper seeks to understand if existing geospatial modelling of onshore wind can account for the low levels of acceptance of onshore wind in the UK (Figure \@ref(fig:acceptanceRates)). In addition, existing social and demographic parameters which can be used to enhance acceptance rate prediction for the site. 

# Background 

## Geospatial Modelling

To assist in the development of onshore wind energy, many methodologies have been produced to determine site suitability for wind farms. Development primarily started within the late 1990s [@Voivontas1998; @Baban2001], and established a structure which has been applied extensively internationally [@Hansen2005; @Yue2006; @Lee2009; @Janke2010; @SQWEnergy2010; @Aydin2010; @VanHaaren2011;  @Sliz-Szkliniarz2011; @Gass2013; @Neufville2013; @Miller2014; @Wang2014; @Watson2015; @Noorollahi2015; @Atici2015; @Baseer2017]. These methodologies combine geospatial modelling with Multi-criteria Decision Analysis (MCDA) techniques to identify sites which are deemed suitable for development [@Malczewski2004].

When determining suitable sites for development, ideal sites are typically identified as 1) *having high average wind speeds*; *2) not being close to urban areas*; *3) not in protected landscapes (e.g. National Parks)*; *4) not close to airports (to minimise radar interference)*; *5) close to roads for access* and *6) close to power lines for grid connection*. However, there are concerns that geospatial parameters in isolation are in themselves insufficient to explain patterns of development of wind turbines [@VanderHorst2010]. These concerns can be further supported by the continued reductions in the acceptance of wind turbine projects in the UK, shown in Figure \@ref(fig:acceptanceRatesWind), suggesting there is a widening gap between existing modelling approaches and real world development patterns.

```{r acceptanceRatesWind, message=FALSE, warning=FALSE, fig.cap = "Annual Average Acceptance rate of wind turbine projects within the UK", cache = TRUE}
turbines <- 
  read.csv(file = "_data/Public_Database_-_Feb_2018.csv", check.names = TRUE, as.is = TRUE) %>%
  mutate(date = lubridate::dmy(`Planning.Application.Submitted`),
         year = lubridate::year(date)) %>%
  filter(Country != "Northern Ireland") %>%
  merge(planning_summary, by.x = "Development.Status..short.", by.y = "Status") %>%
  dplyr::group_by(year, Status.Summary) %>%
  dplyr::summarise(sum = n()) %>%
  group_by(year) %>%
  mutate(freq = sum / sum(sum)) %>%
  filter(Status.Summary == "Approved") %>%
  as.data.frame()

ggplot(turbines, aes(x = year, y = freq)) +
  geom_smooth(method = "lm") +
  geom_point() +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) +
  scale_x_continuous(breaks = seq(1990, 2018, by = 4)) +
  labs(x = "Year", y = "Acceptance Rates (%)") +
  theme(aspect.ratio = 1)
```

Whilst some studies compare the resulting suitability maps with locations of operational wind turbines [@Watson2015; @Miller2014; ; @Gass2013; @VanHaaren2011; @Aydin2010], these were largely used only as a form of discussion, and the information was not directly used to develop or revise the models. This overlooks a valuable contribution that existing sites could provide in understanding whether there are any spatial development patterns which can be identified. In particular, Watson [-@Watson2015] noted "*operational wind farms in South Central England were predominantly located in areas suggested to be of lower suitability*", suggesting that the model inaccurately assesses site suitability in the region.

In situations where there is a large enough sample of similar historical spatial decisions, an "*Inverse theory*" approach can be applied to determine subjective valuation of criteria by stakeholders [@Cirucci2014]. Compared to the traditional "*Forward theory*" approach of geospatial modelling (Figure \@ref(fig:InverseGIS)a), an inverse approach assesses the existing spatial distribution of projects to determine the most influential parameters in determining site success (Figure \@ref(fig:InverseGIS)b). Such an approach has been used successfully within public health studies [@Brody2002; @Mohamed2004; @Yamada2009; @Garcia-Ayllon2013] and infrastructure location decision-making [@USEPA2002; @Cirucci2015] to determine optimal sites for development.

```{r InverseGIS, fig.cap="Comparison of Forward and Inverse GIS MCDA model structures", out.width="50%"}
grViz_pdf("_figures/InverseGIS/InverseGIS.gv")
```

## Planning Acceptance Parameters

The difficulties in receiving planning permission for onshore wind turbines has prompted research to assess the factors which influence the public acceptance of onshore wind turbines. It has been highlighted that the acceptability of wind energy projects is based on more than just geospatial parameters [@Toke2008; @VanderHorst2010; @Langer2016]. In a review of 146 journal articles of planning acceptance, perceptions and attitudes towards wind turbines, Langer et al [@Langer2016] summarised that parameters can be grouped into three groups 1): *Physical and Environmental*; 2) *Pyscho-social* and 3) *Social and Institutional*.

Qualatitive surveys are used to explore potential factors which influence planning acceptance. A key concept consistently investigated within empirical research is the *"proximity hypothesis"*, which states that those living closest to a wind farm will have the most negative perceptions of it [@Devine-Wright2005; @Warren2005]. However, attempts to prove this hypothesis have largely proved unsuccessful, and results have proved conflicting. For example, evidence from Denmark suggests no link between proximity of residential properties to the nearest turbine and negative public perceptions, with suggestions that respondents living closest (i.e. within 500 metres) actually had more positive perceptions in comparison with individuals living away from turbines [@Krohn1999]. This view was further supported by a study in Cornwall, UK, which found that local communities with visibility of the turbines were generally more supportive of wind turbines [@Eltham2008]. However, several studies have reported the opposite relationship [@Meyerhoff2010; @Ladenburg2006], with the studies finding that negative perceptions increased with proximity to wind energy developments.

Literature has also sought to understand the potential cumulative effects that wind turbines have, as projects are frequently refused planning in regions already containing wind turbines [@Strachan2004; @Jones2011; @Eltham2008]. However, there has been limited understanding in how neighbouring wind energy projects may alter the likelihood of nearby wind turbine projects receiving planning, and current research has focussed on smaller case studies [@Jones2011].  
It has been argued within literature that psycho-social factors have become crucial dimensions to explain how local communities interact with, and react to, new wind farm developments [@Langer2016]. The effects of socio-demographic variables on individuals' views of wind farms have also been studied within literature [@Devine-Wright2005; @Warren2010]. Age, gender, experience with wind farms, and use of the land and/or beach were found to be slightly correlated with the attitudes towards wind power in a Danish study dealing with public perceptions of onshore or offshore wind energy projects.

At an individual level, empirical findings suggest that political beliefs are correlated with social acceptance of different low carbon technologies [@Devine-Wright2007]. This is supported by surveys that indicated that only 62% of individuals indicating support for the Conservative party were supportive of new renewable energy developments, compared to 86% and 84% for Labour and Liberal Democrats respectively [@Populus2005].

Studies have highlighted that the interaction of developers with local communities are key indicators of positive planning approval outcomes [@Toke2005; @Devine-Wright2005a; @Wustenhagen2007]. Projects which seek greater approval within their plans are generally more successful than those which are fixed prior to consultation with the local population.

Finally, the ownership structure of a project has been indicated to be a significant influence on the level of public acceptance [@Sonnberger2017; @Haggett2006]. Projects are generally seen as more favourable when owned by local energy cooperatives than by a large energy company or investor with no local connections. This reason has been raised to explain the differences in project success between the United Kingdom and Germany [@Toke2008].

## Quantitative Analysis of Turbine Acceptance

There has been increased use of quantitative analysis to quantify the effect parameters have on the outcome of wind energy planning outcomes. Such approaches build upon the understandings provided within the qualitative analysis explained with Section \@ref(planning-acceptance-parameters), and aim to provide numerical data that can be transformed into usable statistics.

Toke [@Toke2005] conducted logistic regression analysis using data collected for 51 wind energy sites within the UK, and explored how planning outcomes were influenced by the views of key actors within the planning process of wind energy, including local councils, planning authorities and landscape protection groups. The study found that planning acceptance rates were closely associated with the high level of apprehension about such schemes amongst people living in the immediate vicinity, highlighting the importance that social influences have on planning acceptance.

Van der Horst and Toke [-@VanderHorst2010] assessed how local characteristics related to the planning outcome of wind energy projects in England. 117 variables related to education, health, demography, employment and housing were used and compared with the planning outcomes for 77 wind energy projects. Univariate regression analysis was conducted with the Mann-Whitney test being used to analyse the associations between the planning decision outcome and each of the independent variables separately. Several strong associations were identified for planning refusal, including (1) *voter turnout* and (2) *years of potential life lost*^[Years of potential life lost (YPLL) is an estimate of the average years a person would have lived if he or she had not died prematurely. It is, therefore, a measure of premature mortality.]. The study notes that wind energy appears to generally be more likely to receive planning permission in deprived areas, and as previously noted within the review, some developers were "*keen to avoid relatively privileged communities and target areas where people are thought to less likely put up a fight*" [@VanderHorst2010, p.220]. These issues highlight the potential importance of social parameters in site selection.

Van Rensburg et al. [-@VanRensburg20] utilised adjusted probit regression to assess the relative magnitudes of association amongst wind farm project planning approval against a range of 66 variables including project technology, institutional processes and site endowment. Information was collected from 354 wind farm applications and planning authority decisions between 1990 and 2011 in Ireland. The results suggested a range of variables which appeared significant for planning, including 1) *proximity to Natura 2000 sites*; 2) *sites with high bird sensitivity*; 3) *hub height*  and 4) *project capacity*. In addition, the study noted that proximity of the nearest dwellings and wind speeds appeared insignificant, which is counter to the view reported within many previous studies. Of the variables included within the model, it concluded a 0.31 predictive confidence value.

## Combining GIS and Quantitative Research

Whilst studies suggest a relationship between demographic and social data and wind turbine planning acceptance rates, none of geospatial models reviewed attempted to integrate these into their assessment beyond the use of proximity buffers around urban areas. It is argued that this omission fails to fully factor in the social dimension in terms of its impact on the suitability of sites for development.

Existing quantitative studeis highlight the value which can be provided by assessing previous planning applications. However, these projects have been limited in the number of projects considered and the parameters considered. Responding to calls to combine qualitative and quantitative research [@Langer2016, p.256], this paper presents analysis which assesses parameters that influence wind turbine planning outcomes, utilising a range of physical, geographical, demographic and political parameters.


#	Material and Methods

The overall methodology is highlighted in Figure \@ref(fig:Methodology), with a detailed explanation provided in the following subsections.

```{r Methodology, fig.cap="Research methodology"}
knitr::include_graphics("_figures/Flowchart/ResearchApproach.pdf")
```

## Study Scope

The study was conducted across Great Britain (England, Scotland & Wales). These were chosen due to the broadly similar categorisation of land types, nature designation, data availability and legislation across these regions [@HMGovernment2014]. The Shetland Islands were excluded from the analysis as their geographic isolation and distance from mainland Britain created issues in generalising the model results.

### Wind Turbines Dataset

Information for turbine planning applications was collected through the Renewable Energy Planning Database (REPD) [@DECC2016] with planning dates between January 1991 and December 2015 (n=1675). Detailed information for each planning application includes the location; year of application; number of turbines; turbine capacity and planning decision. The planning permission status was summarised to a dichotomous variable for use within the statistical analysis: 1) *Approved* and 2) *Refused/Abandoned*. The spatial distribution of these sites is highlighted in Figure \@ref(fig:StudyExtent).

```{r StudyExtent, fig.asp=1, fig.cap="Location of onshore wind energy planning applications used within the study. Location data extracted from the Renewable Energy Planning Database (REPD)", fig.width=4, message=FALSE, warning=FALSE}

# Load shapefile
basemap <- raster::shapefile("_data/shapefiles/BritainCoastlineNoShetland.shp") %>%
  broom::tidy()

# Add count to facet label
turbines <- invisible(readr::read_csv("_code/outputs/turbine_proximity_full.csv", col_types = readr::cols()))
turbines$facetLabel  <- AddNameStat(turbines, "Status.Summary", "lon", stat = "count")

# Plot results
ggplot() +
  geom_polygon(data = basemap, aes(long, lat, group = group), fill = "grey90", colour = "grey20", size = 0.7) +
  geom_point(data = turbines, aes(X.coordinate, Y.coordinate, fill = facetLabel), 
             shape=21, alpha = 0.8, size = 1) +
  scale_fill_manual(values = c("#7CAE00", "#F8766D")) +
  facet_grid(.~facetLabel, switch="both") +
  coord_equal() +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        legend.position = "none",
        panel.grid = element_line(colour = 'transparent'),
        strip.text = element_text(face = "bold")) 


```

### Model Layers Data

Building upon the literature referenced in Section \@ref(introduction), data sources were identified for geospatial and social parameters which had been indicated to be influence wind turbine planning applications. A summary of the variables is provided in Table \@ref(tab:SummaryTable), with a full details provided within the Technical Appendix A.

```{r SummaryTable, message=FALSE, warning=FALSE}
parameters <- read.csv("_data/ParameterTable.csv", check.names = FALSE, stringsAsFactors = FALSE)

parameters %>%
    select(Category, Variable) %>%
  knitr::kable(booktabs=TRUE,
             caption = "Parameters considered within model",
             format = "latex", 
             linesep = "",
             escape = F) %>%
  kableExtra::footnote(alphabet = c(
    "Roads are broken into four main categories: Motorways, A Roads, B Roads and Minor Roads",
    "High Voltage network at 140 400kV",
    "L4 represents degree level or above",
    "AB represents Higher and intermediate managerial, administrative, professional occupation"), threeparttable = T)
```

- **Resource**: wind speeds were taken from the Numerical Objective Analysis of Boundary Layer (NOABL) wind speed database. This provides estimated annualised wind speed at 45m elevation at a resolution of 1km grid [@DTI2001], and has been used in previous studies within the UK [@Baban2001; @Watson2015]. 
- **Features**: Physical features including roads, railways and urban areas were collected from OS Strategi [@Survey2016]. The electricity transmission network, military sites and airport locations (civil and military) were extracted from Open Street Maps (OSM) [@Overpass2016]. 
- **Landscape and Nature**: Landscape and nature designations were collected for regions within the UK [@Pope2017]. 
- **Geographic**: Site elevation data was collected at a 25m resolution [@Commission2015]. This was used to calculate the gradient using the Fleming and Hoffer algorithm [@Fleming1979].
- **Census**:  Census data was collected at the Lower Super Output Area (LSOA) and Data Zone (Scotland) which represents regions with a population between 1000 and 3000 people [@OfficeforNationalStatistics].  
- **Political**: Political data was collected at the local authority level for the four largest parties in the UK: 1) *Conservatives*; 2) *Labour*, 3) *Liberal Democrat* and 4) - *Scottish National Party (SNP)*. These parties hold a sum  of 95% of council seats within Great Britain as of 2016.
- **Cumulative**: the nearest wind turbine to the project was calculated using the location and year of planning application for each turbine.

It was not possible to collect bird sensitivity maps for the whole study region. Guidances states that this should be considered within the planning of projects [@Gove2013].

- **Bird Mapping**:

##	Data Transformations

The data sources came in a varying spatial features, and were aggregated at each of the turbine locations as follows:

-	**Points, lines and polygons**: A spatial join was completed to find the distance to the nearest feature for each turbine. For polygon based data source, value of 0km denotes the turbine is within the feature. Left censoring was used to limit the maximum distance of geospatial relationships to 30km, preventing extreme values from skewing the datasets.
-	**Tabular**: corresponding political and census boundaries were used to map the tabular data, and turbines assigned the value of the region. In addition, political data was filtered to the year of the planning application to determine the political balance at the time of planning.
-	**Raster**: The raster value at the site location was extracted.

In comparison to previous studies [@VanRensburg20], no transformations were made to the standardise the dataset. Transformation provides no direct benefit to the model other than to allow a direct comparison to be made between the influence of parameters. In addition, model parameters create additional complication in the generalisation of the model results as models must transformed to the adjusted z-score scale to allow for comparison [@Harrell2001, p.123].

##	Statistical Modelling

A multiple logistic regression analysis was conducted to model the factors associated with a positive planning outcome of wind turbine applications using the predictor variables listed in Table \@ref(tab:SummaryTable). This model built upon the approaches developed by Toke et al. [@Toke2005] and Van Rensburg et al. [@VanRensburg20]. A hierarchical approach was applied to the model whereby parameters are added to the model sequentially based on the presumed importance of parameters. These were selected as follows:

1. **Aspatial Site Attributes**: variables including *Number of Turbines* and *Installed Capacity*.
2. **Economic Considerations**: parameters which influence the cost effectiveness of the site, including *Wind Speed* and *Proximity to the National Grid*.
3. **Temporal Aspect**: the year in which the planning application was made.
4. **Proximity to Features**: proximity to geographic features, Landscape and Nature Designations
5. **Social and Census Data**: Demographic data for the area of the wind energy project
6. **Political Data**: the political composition of the local authority composition at the time of the planning application.
7. **Spatial Proximity to Other Turbines**: the proximity to the nearest wind energy project.

For each additional set of parameters added to the model, diagnostic checks were made to ensure that the assumptions of logistic regression were maintained. Each parameter was checked for linearity of the logit for independent variables, absence of multicollinearity and independence of variables [@Harrell2001]. Any parameters which violated these conditions were removed from the model.  The overall fit of the model was assessed using Pearson chi-squared, Psuedo R^2^ values and the residual deviance. Internal validation was used to assess the predictive accuracy of the model, with a random sample of 5% fold size randomly selected and iterated 200 times. Once all parameters had been included within the model, a parsimonious model was produced to remove uninfluential parameters, with the Akaike Information Criterion (AIC) used to determine the best fitting model.

## Regional Variation

Regional differences in parameters effects between England, Scotland and Wales were hypothesised due to differing population densities (England: 413/km^2^, Wales: 149/km^2^, Scotland: 68/km^2^) [@ONS2013] as well as differing institutional support, with Scotland in particular placing a greater emphasis on the development of onshore wind [@DECC2016]. Where significant variation is expected within subgroups, split data models are recommended, whereby the dataset was segmented into groups based on model variables, and regression models are fitted to each subgroup [@Stoltzfus2011].^[Whilst a dummy variable could also be considered, such an approach only captures the *level* effect and not the *slope* effect, and therefore only allows for limited variability between the subgroups.] Separate logistic regression models were produced, with the dataset split for each country. The parameters used to construct these models used two approaches. Firstly, the parameters from the parsimonious model were used across all three models ("*Global Parameter*"). In addition, an all-subset regression approach was used within each group to identify the best-fitting model within each group ("*Optimised Parameters*").

## Generalisation

A spatial regression models was used to generalise the findings of statistical analysis, and are frequently used within geospatial modelling [@Ward2008]. The outputs of the parsimonious regression model were used to generalise the results for national prediction. This model contained 13 variables, two of which were non-spatial parameters, *Turbine Capacity* and *Year*. To include these within the prediction, fixed values were assumed, and predictions made for the year 2017 with a turbine size of 2MW (the average model size for the given year [@DECC2016]).

#	Results 

The overall results for each stage of the hierarchical model are presented in Table \@ref(tab:LogisticModelComparison). It can be seen that there is a marginal improvement of the Nagelkirke R^2^ values across the results, and similarly the predictive accuracy of the model improves as more parameters are included. The final model reports a Hosmer-Lemeshow p-value of 0.032, which indicates that the null hypothesis holds for this model.

```{r LogisticModelComparison}

knitr::kable(model_results,
             format = "latex",
             caption = "A summary of the hierarchical logistic regression models",
             booktabs=TRUE)  %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "scale_down"))
```

Table \@ref(tab:LogisticResults) provides the results from the final hierachical regression model (Model 7). Statistically significant positive trends (e.g. increase in the parameter increases success rates) were observed for 1) *Turbine Capacity*; 2) *Distance to Urban Regions*; 3) *Distance to AONB* 4) *Distance to National Parks* and 4) *Distance to Nearest Turbine (Rejected)*. Negative associations were found for 1) *Year*; 2) *Distance to Ramsar sites* 3) *Distance to Natura 2000 sites*; 4) *Qualifications L4 (University degree or above)*; 5) *Mean Age*; 6) *Nearest Turbine (Operational)*.

The total number of parameters retained in the parsimonious model was reduced to 15. This resulted in a marginal penalty in performance of the model, with the R^2^ values reducing from 0.2 to 0.193. The odds ratios (OR) for these remaining parameters are shown for each parameter in Figure \@ref(fig:oddsPars), whereby an OR = 1 means the parameter does not affect odds of the planning outcome, OR > 1 indicate the parameters positively influence planning acceptance, OR < 1 represents a negative parameter influence.


```{r LogisticResults}
knitr::kable(fullResultsTable,
             booktabs =TRUE,
             caption = "Odds Table for Logistic Regression Parameter") %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "scale_down"))
```


```{r oddsPars, fig.cap="Odds Plot for Parsimonious Logistic Regression Model", out.width= "100%", fig.asp=0.7}
OddsPlotPars
```

The results of the nationally-segmented models are summarised in Table \@ref(tab:kableNestedModelResults), comparing the *Global Parameters* against *Optimised Parameters*. There has been a general increase in the fit of the models represented by the Nagelkirke R^2^ values. The difference between Odds Ratios between each model is further highlighted in Figure \@ref(fig:OddsPlotSegmented).

```{r kableNestedModelResults, warning=FALSE}
summary_split_model %>%
  knitr::kable(caption = "Comparison of subset Logistic Regression Models based on the global parameters list", format = "latex",
             booktabs=TRUE) %>%
  kableExtra::kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  kableExtra::add_header_above(c(" " = 2, "Global Parameter" = 3, "Optimised Parameters" = 3))
```

(ref:OddsPlotSegmented) Odds plot for the nested logistic regression models for England, Scotland and Wales. Error bars indicate 95% confidence intervals.

```{r OddsPlotSegmented, fig.cap= "(ref:OddsPlotSegmented)", out.width= "100%", fig.asp=0.7}
segmented_country$plot +
  theme(aspect.ratio = 1) +
  scale_fill_manual(values = c("#70AFD1", "#CBE1EE", "#B8C7CE"))
```


```{r, cache = FALSE}
r <- raster::raster("_code/outputs/PredictedRaster.tif")

average <- raster::cellStats(r, stat='mean') %>%
  scales::percent()

# Above 50% acceptance
above50 <- raster::cellStats(r>0.5, stat='mean') %>%
  scales::percent()

# Above 50% acceptance
below10 <- raster::cellStats(r<0.1, stat='mean') %>%
  scales::percent()
```

Finally, the results of the statistical generalisation are presented in Figure \@ref(fig:predictionRaster). The average acceptance value of the model is `r average`. Only `r above50` of sites were predicted to have a greater than 50% chance of success, and `r below10` of sites were predicted to have an acceptance rate of less than 10%.

```{r predictionRaster, fig.cap="Predicted Raster of Site Acceptance", fig.asp= 1.5}
r <- raster::raster("_code/outputs/PredictedRaster.tif")
rdf <- raster::rasterToPoints(r) %>%
  as.data.frame()

# Define the cut points used within the plot
cutpoints <- 
  function(x) cut(x, breaks=seq(0,1,0.1), labels=c(" 0-10","10-20","20-30","30-40", "40-50", "50-60", "60-70", "70-80", "80-90", "90-100"))

# Assign Continuous value into bins
rdf$Bin <- cutpoints(rdf$PredictedRaster)

# Plot results
ggplot(rdf, aes(x=x, y=y, fill=Bin)) + 
  geom_raster() + 
  coord_equal() +
  labs(x = "", y = "", fill = "Score") +
  scale_fill_manual( values = c("#a50026", "#d73027", "#f46d43", "#fdae61", "#fee08b", "#d9ef8b", "#a6d96a", "#66bd63", "#1a9850", "#006837")) +
    theme(legend.position = "right",
        axis.line=element_blank(),
        legend.direction="vertical",
        legend.key.height=unit(1,"line"),
        axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank())


```


#	Discussion

The overall model fit of parameters is comparatively low based on geospatial parameters alone with the global parsimonious model achieving a fit of 0.2. Compared to Van Rensburg et al. [@VanRensburg20], an overall McFadden adjust R^2^ value of 0.31 was determined. Although fewer geospatial parameters were included within the former analysis, the study integrated greater institutional details and planning details.

## Significant Parameters

For project characteristics, the size of the turbine capacity is indicated as a significant parameter, with larger turbines increasing the chance of acceptance. This at first appears counter intuitive, but may suggest that the developers of bigger turbines are more likely to appeal the decisions made against larger wind farms, as rejection of such projects would result in a large loss of potential revenue. However, it should be noted that this variable has a small standard deviation (sd = 1.0) compared to other variables included within this model, and therefore the odds ratio inflates their influence within the model.

The distance to urban areas was indicated to be statistically significant, although there is considerable uncertainty as indiciated by the confidence interval. There are a number of potential causes for this: firstly, it could indicate that high wind speed sites suitable for development tend to be naturally less populated (i.e. hilly, isolated regions). Additionally, it may reflect a so-called "Not in My Back Yard" (NIMBY) view from the vocal local population, with projects in closer proximity to urban areas being more likely to be rejected. This has been a relatively contentious subject within literature, with a range of studies supporting [@Haggett2006; @Jones2010a] and rejecting [@VanRensburg20; @Devine-Wright2005a; @Populus2005] the NIMBY argument. However this study provides quantitative evidence to suggest that that sites closer to urban areas have a lower chance of acceptance. 

For landscape and environmental designations, distance to National Parks, Ramsar and AONB were indicated as significant parameters although have marginal impacts. This potentially reflects the negative visual impacts which are often cited as a major impact of wind energy developments [@Langer2016; @Jones2010a]. However, it should be noted that these influences have a relatively low impact, despite literature indicating that landscape designations would play a more important role.

The level of qualifications, and the mean age of the local populous have been retained as significant parameters for demographic variables. It is suggested that regions of higher education may be more effective in organising campaign groups against such projects. This supports the hypothesis developed by Van der Horst and Toke [@VanderHorst2010] that developers were that developments are likely to avoid more privileged areas. To the author's knowledge, such a connection between acceptance rates and demographics has not be previously quantitatively validated.

The analysis suggests that proximity to existing wind energy developments may influence the likelihood of projects receiving planning. The nearest operational wind energy project was indicated as having a statistically significant negative effect, which suggests that projects further away from an existing project are less likely to be accepted. In addition, the nearest rejected project is suggested to be have a negative estimate, inferring that the further the site is from a previously rejected project, the higher the chance of acceptance. This "proximity hypothesis" has been a contentious subject challenged within literature [@Meyerhoff2010; @Ladenburg2006; @Eltham2008]. However, this study provides quantitative evidence to challenge this view.

There are notable parameters which are frequently used in GIS modelling, but do not prove influential, including wind speed and the proximity to airports. This may reflect that these parameters represent technical challenges which can be investigated in the early stages of project development, and therefore any sites that are not suitable will not seek planning permission.

## National Models

The split data model developed suggests that despite hypthosized differences between the three countries, there is limite variation between the significant influential parameters as shown in Figure \@ref(fig:OddsPlotSegmented). Only *Wind Speed* exhibited statistically significant variations influences, with the model suggesting that sites in England have a greater chance of acceptance.  However, the reduced number of observations used to build each model increases the uncertainty substantially as indicated by the confidence intervals. 

## Generalisation

The national prediction raster shown in Figure \@ref(fig:predictionRaster) highlights that there are large regional variations within wind energy site acceptability. For example, large regions in Scotland appear suitable for development, while many other regions with the South of England appear "off limits" to development, particularly the regions along the South Coast of the UK. 

For the overall model, there is a low average predicted acceptance rate of `r average` which is below the rate of acceptance of wind energy in the UK, which was 40% in 2017 [@DECC2016]. It would be expected that the model would return a lower average, as sites which are selected by developers will pass through several preselection criteria prior to planning permission [@Smith2016]. Therefore, sites which are generally opposed before planning will often be abandoned before being taken to planning permission. 

The analysis results suggest that there is no "one-size-fits-all" approach for spatial modelling, and that there are large regional variations in the development of wind energy projects beyond the availability of the resource. In comparison, the regional renewable energy studies conducted within the UK in 2010 broadly followed a consistent methodology to assess the resource potential, with small differences in the development rules in particular regard to environmental and landscape designations [@Stoddart2012]. It is therefore important that geospatial modelling aims to integrate local understanding to more accurately capture this variation.
 
Surprisingly, the model suggests that the South West of England has a low likelihood of acceptance, despite having high levels of wind energy within the area. The UK's wind energy development largely started in the region, so it had generally been considered supportive of wind energy [@Eltham2008]. Upon inspection

#	Conclusions & Policy Recommendations

This paper has investigated the influence of geospatial, environmental, demographic and political attributes on the probability of wind farm planning approval in Great Britain between 1990 and 2016. The study findings reveal that local demographic parameters appear to influence the planning outcomes of projects, and that many of the geospatial parameters typically integrated into wind turbine models appear insignificant in determining site approval. To the authors' knowledge, such quantitative findings have not previously been demonstrated using such datasets.

UK energy policy has shifted towards a more hostile stance against onshore wind energy since 2015. In particular, the approval of planning has been granted to local communities.  It appears that certain demographics are less accepting of onshore wind in Great Britain. Given that UK planning policy has now devolved power locally and allowing local communities to have the final say on projects [@Smith2016], there may be a clear block to development in certain regions in the country.



In addition, the results raise concerns of the predictive ability of existing geospatial modelling in locating wind energy sites. These findings provide evidence to support existing literature that GIS tools in themselves are of limited applicability [@Toke2005; @Malczewski2004], and supports the conclusion that greater emphasis needs to be given to the non-physical elements of a project (e.g. Community engagement with the scheme from an early stage) [@Toke2008; @Wolsink2000; @Warren2010].

Although the results have been calculated within the context of the UK, these findings are of interest internationally. There are opportunities for spatial plannings to acknowledge the demographic influence on wind energy development. It


In relation to future UK onshore wind, non-technical parameters (such as voter preference, social opinion) may change, and therefore the outcomes of these results may help inform the location of socially acceptable wind turbine sites.

It should be noted that the parameters used to derive these findings are obtained with context to Great Britain, and therefore may have limited applicability internationally, and therefore should be applied with caution outside of this region. There are opportunities to expand upon this work by exploring the international context of the finding to widen its applicability.

The generalised map highlight that there are regions in the UK which still appear

However, with the estimated cost of planning applications for commercial scale projects exceeding \pounds 50000 [@RF2016], there is large value in even marginal improvements in the site selection. The findings from this model can help inform regional level strategy and provide an insight to developers of where projects may be more suitable for future development

POLICY

- Uk transferred planning permission to local communities.



<!--- Ending new subsidies for onshore wind 
There are signs that the growing policy is being reconsidered


However, a clear understanding of the social issues surrounding onshore wind development has not been formulated properlu.



The current policy is both 

http://data.parliament.uk/writtenevidence/committeeevidence.svc/evidencedocument/business-energy-and-industrial-strategy-committee/clean-growth-strategy/oral/74989.html

http://www.alphr.com/energy/1008375/uk-renewable-energy-progress-2020
https://www.gov.uk/government/speeches/ending-new-subsidies-for-onshore-wind
--->

There is recent evidence that the UK government may be looking to soften the stance against onshore wind energy http://data.parliament.uk/writtenevidence/committeeevidence.svc/evidencedocument/business-energy-and-industrial-strategy-committee/clean-growth-strategy/oral/74989.html

## Acknowledgments {-}

This work is part of the activities of the Energy and Climate Change Division and the Sustainable Energy Research Group at the University of Southampton [www.energy.soton.ac.uk](www.energy.soton.ac.uk). It is also supported by ESPRC under grant EP/J017698/1, Transforming the Engineering of Cities to Deliver Societal and Planetary Wellbeing and the Faculty of Engineering and Environment at the University of Southampton. 

## Supplementary Files {-}

```{r include=FALSE}
knitr::write_bib(x = c("knitr", "rmarkdown", "ggplot2", "raster"), file = "software.bib")
```

 The analysis and report was written using the R programming language and RMarkdown [@R-rmarkdown]. The full statistical analysis and turbine dataset is provided with the supporting files available to recreate the results presented in this report.

A online appendix is provided which provides the full statistical analysis and diagnostic checks.

# References {-}

<div id="refs"></div>

Add reference: [@Scherhaufer2017]

\newpage
# Appendix {-}

```{r message=FALSE, warning=FALSE}
knitr::kable(read.csv("_data/ParameterTable.csv", check.names = FALSE),
             booktabs=TRUE,
             caption = "Summary of data sources used within model",
             format = "latex",
             escape = F)  %>%
    kableExtra::landscape() %>%
  kableExtra::kable_styling(latex_options = "scale_down") %>%
  kableExtra::footnote(alphabet = c(
    "Roads are broken into four main categories: Motorways, A Roads, B Roads and Minor Roads",
    "High Voltage network at 140 400kV",
    "L4 represents degree level or above",
    "AB represents Higher and intermediate managerial, administrative, professional occupation"))
```

